Cluster Difference Imaging Photometric Survey. III. An Open Cluster With a 300\,pc Halo

### 2021-03-06 cluster_paper/ms.tex
update ms

\newcommand{\numcdipslcs}{671{,}901\ } % drivers.yield_calculation.count_cdips_lcs.py; 2021.01.13
\newcommand{\numsouthernuniqlcs}{483{,}407\ } % drivers.yield_calculation.count_cdips_lcs.py; 2021.01.13

\newcommand{\nkinematic}{3{,}298\ } % "fullfaint" kinematic sample, core+halo (CG18+KC19+M21). from any call of earhat.helpers._get_fullfaint_dataframes
\newcommand{\nnbhd}{13{,}843\ } % make_compstar_NGC_2516_sourcelist.py
\newcommand{\ncore}{1{,}106}  % "fullfaint" kinematic sample, CG18. from any call of earhat.helpers._get_fullfaint_dataframes
\newcommand{\nhalo}{2{,}2192} % "fullfaint" kinematic sample, KC19+M21. from any call of earhat.helpers._get_fullfaint_dataframes

% from drivers.cluster_rotation.get_auto_rotation_periods.py
% checked with plot_target_vs_actual_list_sanity_checks.py
\newcommand{\ncdips}{2{,}205\ }
\newcommand{\nkinematicrpltsixteen}{2{,}270\ }
\newcommand{\nkinematicCGandKC}{2{,}236\ }

@@

%FIXME these change depending on your cutoff. ensure the final numbers agree!
\newcommand{\nautorotdenominator}{1{,}641\ } % plot_autorotation.py, w/ runid=NGC_2516. For nequal<=0, nclose<=1, LSP>0.08. See /doc/20210205_nequal_and_Prot_vs_color_cleanup.txt
\newcommand{\nautorotnumerator}{987\ } % plot_autorotation.py, For nequal<=0, nclose<=1, LSP>0.08. See /doc/20210205_nequal_and_Prot_vs_color_cleanup.txt
\newcommand{\ncompstardenominator}{820\ } % plot_autorotation.py, w/ runid=compstar_NGC_2516. For nequal<=0, nclose<=1, LSP>0.08. 
\newcommand{\ncompstarnumerator}{365\ } % plot_autorotation.py
\newcommand{\ncompfrac}{44.5\%} % plot_autorotation.py
\newcommand{\nautovscompstardenominator}{705\ } % plot_autorotation.py, w/ runid=compstar_NGC_2516. For nequal<=0, nclose<=1, LSP>0.08. 
\newcommand{\nautovscompstarnumerator}{374\ } % plot_autorotation.py
\newcommand{\nautofrac}{53.1\%} % plot_autorotation.py

@@

In the right panel of Figure~\ref{fig:hr}, we compare the observed
Gaia EDR3 photometry with PARSEC isochrones
\citep{bressan_parsec_2012,chen_improving_2014,chen_parsec_2015,marigo_new_2017}.
We used the web
interface\footnote{\url{http://stev.oapd.inaf.it/cgi-bin/cmd},
\texttt{2021-02-26}, \texttt{CMD3.4}, \texttt{YBC} bolometric
corrections as in \citet{chen_2019_YBC}.} to interpolate these
isochrones at $\log_{10}(t/\mathrm{yr})=\{8, 8.25, 8.5\}$, and applied
the reddening correction described in Section~\ref{subsec:tess}.  We
assumed a reddening of $E(B-V)=0.102\pm0.019$, $R_V=3.1$, and applied
the \citet{odonnell_1994} SED-dependent extinction law star-by-star
through the PARSEC interface.  For the metallicity, we considered a
range of super and sub-solar metallicities to fit as much of the locus
from $0.5<Bp-Rp<1.5$ as possible, and settled on the slightly
super-solar $[M/H]=0.06$ \citep{cummings_2011_li_iron}.  Sub-solar
metallicities led to model predictions too blue along the main
sequence by $\approx$0.1\,mag.  Our adopted metallicity agrees with
the spectroscopic metallicity from
\citet[][Sec~4.4.4]{cummings_2011_li_iron}, and with the iron
abundance recently determined by the Gaia-ESO team
\citep{baratella_gaiaeso_2020}, which represented an up-revision from
an earlier sub-solar metallicity \citep{randich_gaiaeso_2018}.  While
\citet{bailey_rv_2018} report a slightly sub-solar metallicity for the
cluster, we prefer super-solar based on the goodness of fit from the
photometry.

@@

\begin{figure*}[t]
	\begin{center}
		\leavevmode
    \includegraphics[width=0.75\textwidth]{f3a.pdf}
		% \subfloat{
		% 	\includegraphics[width=0.75\textwidth]{f3a.pdf}
		% }

		% \subfloat{
		% 	\includegraphics[width=0.75\textwidth]{f3b.pdf}
		% }
	\end{center}
	\vspace{-0.7cm}
	\caption{ {\bf The core and halo of \cn\ in the space of rotation
    period and dereddened Gaia color.}
    The Pleiades \citep[125\,Myr][]{rebull_rotation_2016a}
    and Praesepe \citep[650\,Myr][]{douglas_poking_2017}
    are shown for reference.
    Stars in the core and the halo of NGC\,2516 appear
    gyrochronologically coeval with the Pleiades at (Bp-Rp)$_0$$<$1.25.
    There is a narrow range of color, from 1.25$<$(Bp-Rp)$_0$$<$1.75,
    in which some NGC\,2516 members appear elevated with respect to
    the Pleiades, suggesting a slightly older age.
		\label{fig:rot}
	}
\end{figure*}

@@

We then measured the rotation period of the resulting light curve.  We
used the CDIPS aperture radius that, based on theoretical expectations,
was expected to give the optimal balance between light from the target
and background-light \citep{Sullivan_2015}.  This typically resulted in an
aperture radius of either 1 or 1.5 pixels.  To measure the periods, we
used the periodogram implementations in \texttt{astrobase}, in
particular the Stellingwerf PDM periodogram (CITE), along with the
more traditional Lomb-Scargle (CITE).  We recorded the top five
periodogram peaks from each method, and their corresponding powers.
Finally, as a check on crowding, we also recorded the number of stars
within the aperture of equal brightness to the target stars, and of
brightness with 1.25 and 2.5 TESS magnitudes of the target star.

Figure~\ref{fig:rot} shows the resulting rotation periods.  The points
on this plot are the \nautorotnumerator light curves for which the peak Lomb Scargle
periodogram period was below 15 days, the LS power exceeded 0.08, and
for which no equal-brightness or greater companions were within the aperture. 
We also required that at most one companion with brightness exceeding 
one tenth of the target star could reside in the aperture.
These
selection criteria are entirely heuristic, and yielded
\nautorotdenominator stars for which rotation periods might have been
detected.  
By exploring the different possible selection criteria, we found that
the ones we adopted provided a
reasonable balance between excluding stars on the slow sequence, {\it
vs.} reporting spurious rotation periods in the ``void'' below the
slow sequence, where the dominant signal was in fact from a
neighboring star.
Splitting the resulting sample into the ``core'' and the ``halo'', they
suggest that the halo is real.

To put the clusters on the same color scale, we needed to adopt
reddening corrections.
% FIXME: where did these reference tables come from???
We assumed {X} for the Pleiades, and {Y} for Praesepe.
For NGC\,2516, we used the \texttt{astrobase.services.dust} module to
query the 2MASS DUST
service\footnote{\url{http://irsa.ipac.caltech.edu/applications/DUST}}
and retrieve the extinction parameters for the positions of each
NGC\,2516 member \citep{bhatti_astrobase_2018}.  The mean and standard
deviation of the $E(B-V)$ values for the \citet{schlegel_maps_1998}
and \citet{schlafly_measuring_2011} maps were consistent, so we
adopted the average from \citet{schlegel_maps_1998}:
$E(B-V)=0.206\pm0.039$.  \citet{bonifacio_search_2000}
\citet{schlegel_maps_1998} maps overestimate the reddening values when
the color excess exceeds about 0.10 mag. Applying the correction
proposed by \citet{bonifacio_search_2000} leads to our adopted
$E(B-V)=0.102\pm0.019$.  Finally, to convert to Gaia colors we adopted
the calibration of \citet{stassun_TIC8_2019}, namely $E(Bp-Rp)=1.31
E(B-V)$ and $A_G=2.72 E(B-V)$.  This yielded $E(Bp-Rp)=0.134\pm0.025$,
which is used in the plots.

The resulting gyrochronology age we find for the core is XXX.
For the halo, the claimed age from gyrochronology is YYY.
Applying the same procedure to the field star comparison sample (see
Appendix~\ref{app:compstar}),
we get an age of ZZZ+/-AAA.

>>>>

% gaia target sample numbers
\newcommand{\nkinematic}{3{,}298\ } % "fullfaint" kinematic sample, core+halo (CG18+KC19+M21). from any call of earhat.helpers._get_fullfaint_dataframes
\newcommand{\nnbhd}{13{,}843\ } % make_compstar_NGC_2516_sourcelist.py
\newcommand{\ncore}{1{,}106}  % "fullfaint" kinematic sample, CG18. from any call of earhat.helpers._get_fullfaint_dataframes
\newcommand{\nhalo}{2{,}2192} % "fullfaint" kinematic sample, KC19+M21. from any call of earhat.helpers._get_fullfaint_dataframes

% from drivers.cluster_rotation.get_auto_rotation_periods.py;  checked with plot_target_vs_actual_list_sanity_checks.py
\newcommand{\ncdips}{2{,}205\ }
\newcommand{\nkinematicrpltsixteen}{2{,}270\ }
\newcommand{\nkinematicCGandKC}{2{,}236\ }

% cdips results numbers
\newcommand{\numcdipslcs}{671{,}901\ } % drivers.yield_calculation.count_cdips_lcs.py; 2021.01.13
\newcommand{\numsouthernuniqlcs}{483{,}407\ } % drivers.yield_calculation.count_cdips_lcs.py; 2021.01.13

@@

\newcommand{\nautorotdenominator}{1{,}641\ } % plot_autorotation.py, w/ runid=NGC_2516. For nequal<=0, nclose<=1, LSP>0.08. See /doc/20210205_nequal_and_Prot_vs_color_cleanup.txt
\newcommand{\nautorotnumerator}{987\ } % plot_autorotation.py, For nequal<=0, nclose<=1, LSP>0.08. See /doc/20210205_nequal_and_Prot_vs_color_cleanup.txt
\newcommand{\nautorotnumeratormatching}{507\ } % plot_autorotation.py, LSP=SPDM period
\newcommand{\ncompstardenominator}{820\ } % plot_autorotation.py, w/ runid=compstar_NGC_2516. For nequal<=0, nclose<=1, LSP>0.08. 
\newcommand{\ncompstarnumerator}{365\ } % plot_autorotation.py
\newcommand{\ncompfrac}{44.5\%} % plot_autorotation.py
\newcommand{\nautovscompstardenominator}{705\ } % plot_autorotation.py, w/ runid=compstar_NGC_2516. For nequal<=0, nclose<=1, LSP>0.08. 
\newcommand{\nautovscompstarnumerator}{374\ } % plot_autorotation.py
\newcommand{\nautofrac}{53.1\%} % plot_autorotation.py

@@

In the right panel of Figure~\ref{fig:hr}, we compare the observed
Gaia EDR3 photometry with PARSEC isochrones
\citep{bressan_parsec_2012,chen_improving_2014,chen_parsec_2015,marigo_new_2017}.
We used the web
interface\footnote{\url{http://stev.oapd.inaf.it/cgi-bin/cmd},
\texttt{2021-02-26}, \texttt{CMD3.4}, \texttt{YBC} bolometric
corrections as in \citet{chen_2019_YBC}.} to interpolate these
isochrones at $\log_{10}(t/\mathrm{yr})=\{8, 8.25, 8.5\}$.

To determine the reddening correction across the NGC\,2516 cluster, we
queried the 2MASS
DUST
service\footnote{
	\url{http://irsa.ipac.caltech.edu/applications/DUST};
	query performed using the \texttt{astrobase.services.dust} module
	\citep{bhatti_astrobase_2018}.
}
and retrieved the extinction parameters at the positions of each
NGC\,2516 member.  The mean and standard
deviation of the $E(B-V)$ values for the \citet{schlegel_maps_1998}
and \citet{schlafly_measuring_2011} maps were consistent, so we
adopted the average from \citet{schlegel_maps_1998}: $E(B-V)_{\rm
S}=0.206\pm0.039$.  \citet{bonifacio_search_2000} found however that
the \citet{schlegel_maps_1998} maps overestimate the reddening values
when the color excess exceeds about 0.10\,mag. Applying the correction
proposed by \citet{bonifacio_search_2000} leads to our adopted
$E(B-V)=0.102\pm0.019$.  Finally, to convert to Gaia colors we used
the calibration of \citet{stassun_TIC8_2019}, namely $E(Bp-Rp)=1.31
E(B-V)$ and $A_G=2.72 E(B-V)$.  This yielded $E(Bp-Rp)=0.134\pm0.025$,
which is used in the plots.  To redden the isochrones, we assumed
$R_V=3.1$, and applied the \citet{odonnell_1994} SED-dependent
extinction law star-by-star through the PARSEC interface.

For the metallicity, we considered a range of super and sub-solar
metallicities to fit as much of the locus from $0.5<Bp-Rp<1.5$ as
possible, and settled on the slightly super-solar $[M/H]=0.06$
\citep{cummings_2011_li_iron}.  Sub-solar metallicities led to model
predictions too blue along the main sequence by $\approx$0.1\,mag.
Our adopted metallicity agrees with the spectroscopic metallicity from
\citet[][Sec~4.4.4]{cummings_2011_li_iron}, and with the iron
abundance recently determined by the Gaia-ESO team
\citep{baratella_gaiaeso_2020}, which represented an up-revision from
an earlier sub-solar metallicity \citep{randich_gaiaeso_2018}.  While
\citet{bailey_rv_2018} report a slightly sub-solar metallicity for the
cluster, we prefer super-solar based on the goodness of fit from the
photometry.

@@

\begin{figure*}[t]
	\begin{center}
		\leavevmode
		\subfloat{
			\includegraphics[width=0.49\textwidth]{f3a.pdf}
			\includegraphics[width=0.49\textwidth]{f3b.pdf}
		}
	   %
		%\subfloat{
		%	\includegraphics[width=0.6\textwidth]{f3b.pdf}
		%}
	\end{center}
	\vspace{-0.7cm}
  \caption{ {\bf The core and halo of \cn\ in the space of rotation
    period and dereddened Gaia color.}
    Subset $\mathcal{A}$ ({\it top}) and $\mathcal{B}$ ({\it bottom})
    undergo successive degrees of automated cleaning, described in the
    text.
    The Pleiades \citep[125\,Myr][]{rebull_rotation_2016a} and
    Praesepe \citep[650\,Myr][]{douglas_poking_2017} are shown for
    reference.
    Stars in the core and the halo of NGC\,2516 appear
    gyrochronologically coeval with the Pleiades at (Bp-Rp)$_0$$<$1.25.
    There is a narrow range of color, from 1.25$<$(Bp-Rp)$_0$$<$1.75,
    in which some NGC\,2516 members appear elevated with respect to
    the Pleiades, suggesting a slightly older age.
		\label{fig:rot}
	}
\end{figure*}

@@

We then measured the rotation period from the resulting light curve.
We used the CDIPS aperture radius that, based on theoretical
expectations, was expected to give the optimal balance between light
from the target and background-light \citep{Sullivan_2015}.  This
typically resulted in an aperture radius of either 1 or 1.5 pixels.
To measure the periods, we used the periodogram implementations in
\texttt{astrobase}, in particular the Stellingwerf phase-dispersion
minimization periodogram, along with the more traditional Lomb-Scargle
\citep{lomb_1976,stellingwerf_period_1978,scargle_studies_1982,stellingwerf_period_2011,bhatti_astrobase_2018}.
We recorded the top five periodogram peaks from each method, and their
corresponding powers.  Finally, as a check on crowding, we also
recorded the number of stars within the aperture of equal brightness
to the target stars, and of brightness with 1.25 and 2.5 TESS
magnitudes of the target star.

Figure~\ref{fig:rot} shows the resulting rotation periods, following
three different sets of cleaning criteria.  The first and second sets
of criteria were entirely automated, and yielded subsets we denote
$\mathcal{A}$ and $\mathcal{B}$.  For both subsets, we considered
light curves for which the peak Lomb-Scargle periodogram period was
below 15 days, the normalized Lomb-Scargle power exceeded 0.08, and
for which no equal-brightness or greater companions were within the
aperture.  We also required that the target star be the brightest star
in the aperture, and that at most one companion with brightness
exceeding one tenth of the target star's brightness be present in the
aperture according to the Gaia DR2 source catalog.  For subset
$\mathcal{A}$, this yielded \nautorotnumerator light curves, out of
\nautorotdenominator stars for which rotation periods might plausibly
have been detected.  For subset $\mathcal{B}$, we additionally
required that the Lomb-Scargle and Stellingwerf phase-dispersion
methods yielded periods identical to within 10\%.  This yielded
\nautorotnumeratormatching light curves.
Subset $\mathcal{A}$ is more complete than 
subset $\mathcal{B}$, at the expense of greater contamination.

{\bf TODO CORRECTLY}
To put the clusters on the same color scale, we needed to adopt
reddening corrections.
For \cn, we used the same average reddening as described in
Section~\ref{subsec:hr}.
% FIXME: where did these reference tables come from???
For the Pleiades, and for Praesepe, we used {\bf XXX, YYY}.

{\bf TODO CORRECTLY}
The resulting gyrochronology age we find for the core is XXX.
For the halo, the claimed age from gyrochronology is YYY.
Applying the same procedure to the field star comparison sample (see
Appendix~\ref{app:compstar}),
we get an age of ZZZ+/-AAA.
